Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-23 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
applied john's comments and more edits

\begin{document}
% \title{Transferring Human Spatial Regions from Sensor Data via Analogy}
\title{Representing and Reasoning About Spatial Regions Defined by Context}
% \title{Towards a Cognitive System Than Can Represent and Reason about Spatial Regions Defined by Context}
% \title{Context-Dependent Spatial Regions}
% \title{Qualitative and Analogical Cognition for Spatial Regions Defined by Context}

@@

\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. The identification of context-dependent spatial regions provides a suitable task for studying this integration. Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.}

@@

This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}.

The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy.

The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from the spatial model of an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

\section{From Sensors to Semantics}\label{sec:qsr-gen}

@@

The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e. what can be done there and where. In this work we determine this using the type of objects present in a room and their location, both relative to each other and relative to the room itself. The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.

@@

\from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}

We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding. To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training. For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building. Motivated by the learning and generalization abilities of humans, we have chosen \emph{analogy} as the approach our system will use to solve this problem.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.

We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system. Given two structured representations as input (the base and target), SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target.  Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g. door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}. While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions. While steps have been taken toward incorporating this knowledge  (i.e. the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work. For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b}, and others are able to mediate their behavior when certain functional contexts are detected (e.g. the areas near doors \cite{Zender2008a}).

% Nick: Goal reasoning is a separate issue entirely, and one for a different paper
 % Additionally, existing systems which are able to reason about their behavior are not able to do so with reference to the spatial extent of their goals, and thus would also fail to generate the desired behavior on this count.

There is mounting evidence that analogy, operating over structured qualitative representations can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.  The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics. 
% Therefore, we hope and expect to see more research combining these three fields.

@@

\from{Nick}{TODO!}

>>>>

\begin{document}

\title{Towards a Cognitive System That Can Recognize Spatial Regions Based on Context}
% \title{Representing and Reasoning About Spatial Regions Defined by Context}
% \title{Context-Dependent Spatial Regions}

@@

\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. The identification of context-dependent spatial regions provides a suitable task for studying this integration. 
% Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.
}

@@

This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

\section{From Sensors to Qualitative Representations}\label{sec:qsr-gen}

@@

The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e. what can be done there and where. In this work we implicitly represent context using the types of objects present in a room and their location, both relative to each other and relative to the room itself. The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.

@@

% \from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
% done

We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding. To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training. For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building. We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. 
% This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system. Given base and target representations as input, SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures.  Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required (e.g., for behavior planning or human-robot interaction), topological maps (i.e., graphs of connected regions) are usually built on top of these metric representations using various landmark-detection techniques to segment regions, e.g. door detection \cite{Hawes/etal:2009b} or clustering of perceptual features \cite{Peltason/etal:2009a}. While capable of recognizing regions, these approaches do not consider the semantic information necessary for identifying context-dependent spatial regions. While steps have been taken toward incorporating this knowledge  (i.e. the types of objects present and their arrangement) into spatial representations, current approaches fall short of representing the types of regions discussed in this work. For example, a number of systems attempt to recognize the category of a room by the objects present within it \cite{Hawes/etal:2010b} (largely ignore location), and others are able to mediate their behavior when certain functional contexts are detected (e.g. the areas near doors \cite{Zender2008a}).

There is mounting evidence that analogy, operating over structured qualitative representations, can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.

% The combination of analogy and qualitative representation has been shown to be flexible enough to model problems at different levels of refinement (e.g. edges, shapes, groups) \cite{Lovett&Forbus2011} and to combine semantic as well as geometric information in spatial reasoning tasks (e.g. \cite{Lockwood/etal2008}). Our work demonstrates that the combination of analogy and qualitative spatial representations is useful for problems in cognitive mobile robotics. 
% Therefore, we hope and expect to see more research combining these three fields.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system is able to use analogy to recognise CDSRs in previously unseen environments through a process of analogy with a manually annotated known room. The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with in future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.
